### 2018-01-30 Maths Notes/Statistics/S3/Estimation/Estimation.tex
S3 Notes update

\begin{document}
\begin{center}
\underline{\huge Estimation, confidence intervals and tests}
\end{center}
\section{Estimators}
A statistic that is used to estimate a population parameter is an \textbf{estimator}.\\
A particular individual value is an \textbf{estimate}.\\
\textbf{Bias} is how far the estimator is from the true value\\
\\
If a statistic T is used as an estimator for a population parameter $\theta$ then the bias is:
$$E(T)-\theta$$
If $E(T)=\theta$ then the statistic is unbiased
\subsection{Proving $\mathbf{E(\overline{X})=\mu}$}
\textit{Prove that $\overline{X}$ is an unbiased estimator for $\mu$ when the population is normally distributed}\\
\\
A random sample $X_1,X_2,X_3...X_N$ is taken for a population with $X\sim N(\mu,\sigma^2)$
$$\overline{X}=\frac{1}{n}\sum X$$
$$E(\overline{X})=E\Big(\frac{1}{n}\sum X\Big)$$
$$E(\overline{X})=\frac{1}{n}\big(E(X_1)+E(X_2)+E(X_3)...+E(X_N)\big)$$
$$E(\overline{X})=\frac{1}{n}(\mu+\mu+\mu...+\mu)$$
$$E(\overline{X})=\frac{1}{n}(n\mu)$$
$$E(\overline{X})=\mu$$
\subsection{Variance}
$$Var(\overline{X})=\frac{\sigma^2}{\textrm{Sample size}}$$
\\
We use $S^2$ as an estimator for $\sigma^2$
$$S^2=\frac{1}{n-1}\Big(\Sigma X^2-n\overline{X}^2\Big)$$
\subsection{Uniform Distributions}
\textit{Random variable X is continuously uniform $[0,\alpha]$. A sample $X_1,X_2..X_N$ is taken}\\
\textit{Show that $\overline{X}$ is a biased estimate and state the bias}
$$\overline{X}=\frac{0+\alpha}{2}=\frac{\alpha}{2}$$
$$E(\overline{X})=E(\frac{\alpha}{2})=\frac{1}{2}E(\alpha)=\frac{\alpha}{2}$$
$$\textrm{Bias}=\frac{\alpha}{2}-\alpha=-\frac{\alpha}{2}$$
\newpage
\section{Standard error}
$$\textrm{Standard error}=\frac{\sigma}{\sqrt{n}} \ \textrm{or} \ \frac{s}{\sqrt{n}}$$
\end{document}

>>>>

\begin{document}
\begin{center}
\underline{\huge Estimation, confidence intervals and tests}
\end{center}
\section{Estimators}
A statistic that is used to estimate a population parameter is an \textbf{estimator}.\\
A particular individual value is an \textbf{estimate}.\\
\textbf{Bias} is how far the estimator is from the true value\\
\\
If a statistic T is used as an estimator for a population parameter $\theta$ then the bias is:
$$E(T)-\theta$$
If $E(T)=\theta$ then the statistic is unbiased
\subsection{Proving $\mathbf{E(\overline{X})=\mu}$}
\textit{Prove that $\overline{X}$ is an unbiased estimator for $\mu$ when the population is normally distributed}\\
\\
A random sample $X_1,X_2,X_3...X_N$ is taken for a population with $X\sim N(\mu,\sigma^2)$
$$\overline{X}=\frac{1}{n}\sum X$$
$$E(\overline{X})=E\Big(\frac{1}{n}\sum X\Big)$$
$$E(\overline{X})=\frac{1}{n}\big(E(X_1)+E(X_2)+E(X_3)...+E(X_N)\big)$$
$$E(\overline{X})=\frac{1}{n}(\mu+\mu+\mu...+\mu)$$
$$E(\overline{X})=\frac{1}{n}(n\mu)$$
$$E(\overline{X})=\mu$$
\subsection{Variance}
$$Var(\overline{X})=\frac{\sigma^2}{\textrm{Sample size}}$$
\\
We use $S^2$ as an estimator for $\sigma^2$
$$S^2=\frac{1}{n-1}\Big(\Sigma X^2-n\overline{X}^2\Big)$$
\subsection{Uniform Distributions}
\textit{Random variable X is continuously uniform $[0,\alpha]$. A sample $X_1,X_2..X_N$ is taken}\\
\textit{Show that $\overline{X}$ is a biased estimate and state the bias}
$$\overline{X}=\frac{0+\alpha}{2}=\frac{\alpha}{2}$$
$$E(\overline{X})=E(\frac{\alpha}{2})=\frac{1}{2}E(\alpha)=\frac{\alpha}{2}$$
$$\textrm{Bias}=\frac{\alpha}{2}-\alpha=-\frac{\alpha}{2}$$
\newpage
\section{Standard error}
$$\textrm{Standard error}=\frac{\sigma}{\sqrt{n}} \ \textrm{or} \ \frac{s}{\sqrt{n}}$$
\section{Central limit theorem}
The central limit theorem (C.L.T.) states that if $X_1,X_2,X_3...$ is a random sample, from any distribution with mean $\mu$ and variance $\sigma^2$. Then the sample means, $\overline{X}$ are distributed with a normal distribution.
$$\overline{X}\sim N(\mu,\frac{\sigma^2}{n})$$ 
For this, the sample size must be greater than 50
\end{document}
